---
title: "{{< var tearsheet_title >}}"
subtitle: "Executive Tearsheet · {{< var report_month_year >}}"
format:
  pdf:
    pdf-engine: lualatex
    documentclass: scrartcl
    classoption: ["letterpaper"]
    geometry:
      - left=0.55in
      - right=0.55in
      - top=0.55in
      - bottom=0.55in
    fontsize: 9pt
    fig-align: center
    colorlinks: true
    linkcolor: chnavy
    urlcolor: chnavy
    include-in-header:
      text: |
        \usepackage{xcolor}
        % Lailara Design System v2 palette
        \definecolor{chnavy}{HTML}{1f2e7a}
        \definecolor{chred}{HTML}{cc100a}
        \definecolor{chteal}{HTML}{158f75}
        \definecolor{chcoral}{HTML}{b82d4a}
        \definecolor{chbgpale}{HTML}{d9d9d9}
        \definecolor{chtextmuted}{HTML}{595959}
        \definecolor{chtext}{HTML}{333333}

        % Tight typography for a dense 2-page tearsheet.
        \usepackage{parskip}
        \setlength{\parskip}{2.4pt plus 0.5pt}
        \setlength{\parindent}{0pt}

        % KOMA section heading colors and tighter spacing above/below.
        \usepackage[automark]{scrlayer-scrpage}
        \setkomafont{section}{\color{chnavy}\large\bfseries}
        \setkomafont{subsection}{\color{chnavy}\normalsize\bfseries}
        \setkomafont{disposition}{\bfseries}
        \RedeclareSectionCommand[
          beforeskip=0.6\baselineskip,
          afterskip=0.15\baselineskip
        ]{section}
        \RedeclareSectionCommand[
          beforeskip=0.5\baselineskip,
          afterskip=0.1\baselineskip
        ]{subsection}

        % Title block tightening.
        \setkomafont{title}{\color{chnavy}\Large\bfseries}
        \setkomafont{subtitle}{\color{chtextmuted}\normalsize\itshape}
        \makeatletter
        \renewcommand*{\@maketitle}{%
          \par\vspace{-1.6em}%
          {\usekomafont{title}\@title\par}%
          \ifx\@subtitle\@empty\else
            \vspace{0.2em}%
            {\usekomafont{subtitle}\@subtitle\par}%
          \fi
          \vspace{0.2em}%
          {\color{chnavy}\rule{\linewidth}{0.6pt}}%
          \par\vspace{0.4em}%
        }
        \makeatother

        % Booktabs + table tweaks for the triage list.
        \usepackage{booktabs}
        \usepackage{array}
        \usepackage{colortbl}
        \renewcommand{\arraystretch}{1.06}

        % Compact lists.
        \usepackage{enumitem}
        \setlist{topsep=1pt, partopsep=0pt, itemsep=1pt, parsep=0pt,
                 leftmargin=*, labelsep=0.4em}

        % A "stat" macro for the page-1 lead numbers (red, bold, no break).
        \newcommand{\stat}[1]{\textbf{\color{chred}#1}}
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    out.width: "100%"
---

```{r setup}
#| include: false
suppressPackageStartupMessages({
  library(dplyr); library(knitr); library(scales)
})
source("../R/engagement.R")
eng          <- load_engagement("..")
client_name  <- eng$client_name
client_short <- eng$client_short
is_demo      <- eng$is_demo
smf <- readRDS("../output/frames/sku_master_full.rds")
rrs <- readRDS("../output/frames/retailer_readiness_summary.rds")
cbe <- readRDS("../output/frames/chargebacks_enriched.rds")

n_chargeback_months <- length(unique(cbe$month))
n_skus <- nrow(smf)
total_cb_36mo <- sum(smf$chargeback_total, na.rm = TRUE)
annual_cb <- round(total_cb_36mo * 12 / n_chargeback_months)
total_ttm_rev <- sum(smf$ttm_revenue, na.rm = TRUE)

wm_readiness <- rrs |> filter(retailer == "Walmart") |>
  summarise(n_fail = sum(!overall_pass), n_total = n(),
            pass_rate = mean(overall_pass))
wm_fail <- wm_readiness$n_fail[1]
wm_rev_at_risk <- smf |>
  semi_join(rrs |> filter(retailer == "Walmart", !overall_pass),
            by = "sku") |>
  summarise(rev = sum(ttm_revenue, na.rm = TRUE)) |> pull(rev)
wm_rev_pct <- round(100 * wm_rev_at_risk / total_ttm_rev)

n_gtin_fail <- sum(!smf$gtin_valid, na.rm = TRUE)
n_upc_fail  <- sum(!smf$upc_valid, na.rm = TRUE)
n_both_barcode_valid <- sum(smf$gtin_valid & smf$upc_valid, na.rm = TRUE)
n_dims_missing <- sum(smf$missing_case_dims, na.rm = TRUE)
n_dq_100  <- sum(smf$data_quality_score == 100, na.rm = TRUE)
dq_min    <- min(smf$data_quality_score, na.rm = TRUE)
n_clean_skus <- sum(smf$issue_count == 0, na.rm = TRUE)
dg007        <- smf |> filter(sku == "CHP-DG-007")
dg007_events <- cbe |> filter(sku == "CHP-DG-007") |> nrow()
dsd <- function(x) paste0("$", formatC(round(x), big.mark = ",", format = "d"))
ds2 <- function(x) ifelse(abs(x) >= 1e6, sprintf("$%.1f million", x / 1e6), dsd(x))

ds <- function(x) {
  if (abs(x) >= 1e6) sprintf("%.1f million", x / 1e6)
  else formatC(round(x), big.mark = ",", format = "d")
}

n_retailers <- n_distinct(rrs$retailer)
stage3_skus <- as.integer(n_skus * 5L)
stage3_cb <- round(annual_cb * (stage3_skus / n_skus) * (12 / n_retailers))

total_fix_hours <- sum(smf$est_fix_hours, na.rm = TRUE)

# Fix-effort split, computed so the day-plan hours can't drift from the data.
# Each check-digit correction is ~10 min; each case measurement ~30 min.
n_case_fix  <- sum(smf$missing_case_dims | smf$missing_case_weight, na.rm = TRUE)
hrs_barcode <- (n_gtin_fail + n_upc_fail) * 10 / 60
hrs_case    <- n_case_fix * 30 / 60

data_defect_pct <- cbe |>
  mutate(is_data = !is.na(triggered_by_field) & nzchar(triggered_by_field)) |>
  summarise(pct = round(100 * sum(amount[is_data]) / sum(amount))) |> pull(pct)
data_defect_annual <- round((total_cb_36mo * data_defect_pct / 100) * 12 / n_chargeback_months)
```

```{r draft-watermark}
#| results: asis
cat(eng_watermark_md(eng))
```

## The price of poor housekeeping

\stat{\$`r ds(data_defect_annual)` per year} in retailer chargebacks traced to product data defects (of \$`r ds(annual_cb)` in total chargebacks; the rest is fulfillment), plus stalled-launch revenue loss and shelf loss from deauthorizations linked to data quality.

\stat{\$`r ds(wm_rev_at_risk)` in revenue at risk} at Walmart. `r wm_fail` of `r n_skus` SKUs fail Walmart's required-field check today. Those `r wm_fail` carry `r wm_rev_pct`% of the catalog's trailing twelve-month revenue. Costco and Whole Foods have similar gaps at lower thresholds.

\stat{\$`r ds(stage3_cb)` in projected annual chargebacks (all causes)} at the growth target, if current defect rates hold. Each new SKU and each new retailer adds to the chargeback surface independently. The projection is linear. In practice, defect rates tend to worsen during rapid growth because data entry processes that strain at `r n_skus` SKUs break at `r n_skus * 5`.

## The compliance deadline

GS1 Sunrise 2027 transitions American retail barcodes to 2D codes built on validated GTINs. FSMA Rule 204 makes accurate GTINs a federal food traceability requirement. `r n_gtin_fail` of `r n_skus` SKUs carry GTINs that fail algorithmic validation today. Both deadlines are already in motion. A company that cannot pass a check digit calculation on its barcodes cannot participate in either transition.

## Summary

The defining feature of this problem is the asymmetry between cost and fix. Correcting every data defect that generates chargebacks takes `r round(total_fix_hours)` hours of data entry. The defects have been present for months. The connection between the defects and the charges has never been visible until this audit.

\begin{minipage}[t]{0.32\linewidth}
\centering
\includegraphics[width=\linewidth]{../output/charts/01_chargeback_pareto.png}
\end{minipage}\hfill
\begin{minipage}[t]{0.32\linewidth}
\centering
\includegraphics[width=\linewidth]{../output/charts/06_retailer_readiness_revenue_at_risk.png}
\end{minipage}\hfill
\begin{minipage}[t]{0.32\linewidth}
\centering
\includegraphics[width=\linewidth]{../output/charts/12_growth_projection_chargebacks.png}
\end{minipage}

\newpage

## Top 10 by fix priority

```{r triage-table}
#| results: asis

fmt_dollar <- function(v) {
  ifelse(is.na(v), "—",
   ifelse(v >= 1000,
          paste0("\\$", formatC(round(v), big.mark = ",", format = "d")),
          paste0("\\$", formatC(round(v), big.mark = ",", format = "d"))))
}
fmt_per_hr <- function(v) {
  ifelse(is.na(v), "—",
         paste0("\\$", formatC(round(v), big.mark = ",", format = "d"), "/hr"))
}
fmt_minutes <- function(hr) {
  m <- round(hr * 60)
  ifelse(is.na(m) | m == 0, "—", paste0(m, " min"))
}
# Escape LaTeX special chars in product names (we keep escape=FALSE on
# kable so our pre-formatted dollar strings render correctly).
latex_escape <- function(x) {
  x <- gsub("\\\\", "\\\\textbackslash{}", x)
  x <- gsub("([&%$#_{}])", "\\\\\\1", x)
  x
}

triage <- smf |>
  arrange(desc(savings_per_hour > 0), desc(savings_per_hour)) |>
  slice(1:10) |>
  transmute(
    SKU = sku,
    Product = latex_escape(product_name),
    `Annual cb savings` = fmt_dollar(chargeback_total * 12 / n_chargeback_months),
    `Fix time` = fmt_minutes(est_fix_hours),
    `Savings per hour` = fmt_per_hr(savings_per_hour))

# Header colored navy with white text; rows zebra-striped via colortbl.
cat("\\small\n")
cat("\\rowcolors{2}{chbgpale!40}{white}\n")
cat(kable(triage, format = "latex", booktabs = TRUE,
          escape = FALSE,
          align = c("l", "l", "r", "r", "r"),
          linesep = ""),
    "\n")
```

## Three root causes

**No validation at entry.** The product master has no recorded entry source for any SKU. Nobody knows who entered these records or what process was followed. There is no validation gate between data entry and the live product master.

**The defects concentrate.** `r n_gtin_fail` of `r n_skus` SKUs carry invalid GTIN-14 check digits. `r n_upc_fail` carry invalid UPC check digits. `r n_both_barcode_valid` of `r n_skus` SKUs have both valid barcodes; the `r n_skus - n_both_barcode_valid` that fail carry both invalid GTIN-14 and invalid UPC. Case dimensions are missing for `r n_dims_missing` of `r n_skus`. Data quality scores range from `r dq_min` to 100 — the barcode-defect SKUs score lowest because a wrong check digit is counted as a defect, the case-dimension-only SKUs score next, and the `r n_clean_skus` clean SKUs score 100.

**Margin density does not follow gross revenue.** Walmart is the largest retailer by gross dollars but not the most profitable by net margin percentage. Trade spend is negotiated annually. Chargebacks are generated by data defects that `r client_short` can fix any Tuesday afternoon. Every dollar recovered drops straight to net contribution.

## The crown jewel

CHP-DG-007, Trail Mix Premium, is among the ten highest-revenue SKUs in the catalog (`r ds2(dg007$ttm_revenue)` TTM) and fails every retailer's readiness check — zero of `r n_retailers`. `r dg007_events` chargeback events across `r n_chargeback_months` months, `r dsd(dg007$chargeback_total)` in penalties, `r dg007$deauth_count` store authorizations revoked. It carries three defects: an invalid GTIN-14 check digit, an invalid UPC check digit, and missing case dimensions. Fixing all three takes `r round(dg007$est_fix_hours * 60)` minutes. A `r ds2(dg007$ttm_revenue)` product that fails every retailer it sells to is less than an hour of clerical work from full compliance.

## A one-week turnaround

- **Days 1–2:** Correct the barcode check digits (UPC and GTIN-14) on the `r n_gtin_fail` failing SKUs (~`r round(hrs_barcode)` hours), then measure and enter the missing case dimensions and weights on the `r n_case_fix` SKUs that lack them (~`r round(hrs_case)` hours). About `r round(total_fix_hours)` hours total — that closes the entire data-defect scope behind the chargebacks and the readiness failures.
- **Days 3–4:** Implement a check-digit validation gate on every data entry path. The GS1 algorithm is a single modulo operation — block saves when the digit is wrong.
- **Day 5:** Deploy the Monday Morning Dashboard. Establish a monthly 30-minute data quality review.
- **Result:** The product master supports GS1 Sunrise 2027 compliance, FSMA Rule 204 traceability, and retailer expansion without compounding data debt.

```{r provenance-footer}
#| results: asis
if (!is_demo) cat(eng_provenance_md(eng, "product-data-health-audit", "1.0", ".."))
```
